---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Brad's functions (bfuncs)

This repository is a random smattering of functions I wrote for myself, but you can use them too.

<!-- Need to make help vignettes and link to them below -->

## Programing functions

## Simulation functions

## Data wrangling functions

* mean_center: mean center continuous variables.

## Data analysis functions

* conf_matrix: create a confusion matrix.

* [freq_table: estimate percents and 95 percent confidence intervals](https://rawgit.com/brad-cannell/bfuncs/master/inst/doc/Frequency_Tables.html)

* mean95: calculate mean and 95% confidence interval.

* stats: estimate means and related statistics for continuous variables.

* tabstats: compact table of summary statistics for continuous variables.

## Presentation and dissemination functions

* Add functions for creating tables

* about_data: view number of observations and variables in a data frame, but not any of the data values.

* codebook: create a codebook / data dictionary from a data frame.

## Other functions

-------------------------------------------------------------------------------

#### Installation instructions:
```{r eval = FALSE}
devtools::install_github("brad-cannell/bfuncs")
```


