---
title: "Adding Rcpp files to a package"
author: "Brad Cannell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

I just finished the [optimizing R code with Rcpp](https://www.datacamp.com/courses/optimizing-r-code-with-rcpp) course on DataCamp today. I added a simple C++ file to bfuncs as practice, for future reference, and to make sure I understand the process for using C++ files in R packages (as opposed to interactively). Below are some lessons learned and benchmarks.

[Hadley's chapter](http://r-pkgs.had.co.nz/src.html#cpp) was helpful; however, there were a couple of things that could be more clear.

1. `devtool::use_rcpp()` is depricated. Use `usethis::use_rcpp()` instead.

2. It will tell you to add the tags below "somewhere" in your package. After reading through Stack Overflow, I put them in a stand-alone R script called "rcpp_roxygen_tags.R". It seems to be working.

3. Next, add roxygen headers to the rcpp file using `//` instead of `#`. Do this before building and reloading or documenting.

4. Make sure to add `@export` to the roxygen header or the functions won't be available in the package namespace. In other words, the `// [[Rcpp::export]]` alone does not make the function available to be called by the package.

5. Click buld and reload _before_ documenting.

6. Document (Ctrl/Cmd + Shift + D).

# Example

As an example, I added a simple version of last observation carried forward. This is something I want to work on implementing anyway. Keep in mind that this version only works on numeric vectors and has no error checking at the moment.

Having said that, I'm going to compare the Rcpp function to a home spun R version using loops and to zoo::na.locf.

```{r}
na_locf_r_loop <- function(x) {
  current <- NA
  res <- x
  for (i in seq_along(x)) {
    if (is.na(x[i])) {
      res[i] <- current
    } else {
      current <- x[i]
    }
  }
  res
}
```

## Benchmark

```{r}
set.seed(42)
x <- rnorm(1e5)

# Sprinkle some NA into x
x[sample(1e5, 100)] <- NA
```

Check to make sure we get the same results from all functions

```{r}
all.equal(na_locf_r_loop(x), zoo::na.locf(x))
```

```{r}
all.equal(bfuncs::na_locf(x), zoo::na.locf(x))
```

```{r}
microbenchmark::microbenchmark(
  na_locf_r_loop(x),
  zoo::na.locf(x),
  bfuncs::na_locf(x),
  times = 5
)
```

So, the R loop is very readable, but roughly 46 times slower than the Rcpp loop.

I believe zoo::na.locf is vectorized, but still 25 times slower than the Rcpp function.

The Rcpp function is as readable as the R loop funcion (if you learn how to read a little C++) and much faster.
