---
title: "Using codebook"
author: "Brad Cannell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using codebook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load libraries

```{r}
library(tidyverse)
library(bfuncs)
```

# Load data

By default, `codebook` assumes that you want to make a codebook about a dataset file that you have saved somewhere, as opposed to a data frame you're working on in an r session, but don't intend to save. Therefore, the first thing you will need to do is read the data into the current R session.

For the purposes of making a self-contained example, we will actually just create an example data frame in the code chunk below and use it to demonstrate how to use `codebook`

```{r}
set.seed(123)
df <- tibble(
  id = factor(seq(1001, 1020, 1)),
  gender = sample(c("Female", "Male"), 20, TRUE),
  date = sample(seq.Date(as.Date("2015-09-15"), as.Date("2015-10-26"), "day"), 20, TRUE),
  height = rnorm(20, 71, 10)
) %>% 
  print()
```

## Column types

`codebook` classifies all columns as one of four types and uses these categories to determine which descriptive statistics are given in the codebook document:

1. Categorical with many different categories, for example `id` above     
2. Categorical with few different categories, for example `gender` above     
3. Time, for example `date` above    
4. Numeric, for example `height` above    



