---
title: "Using tabstat with dplyr"
date: "Created: 2017-11-17 <br> Updated: `r Sys.Date()`"
author: "Brad Cannell"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tabstat with dplyr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
Sys.setenv(TZ = "US/Central")
```

```{r load_packages, message=FALSE}
# Load packages
library(tidyverse)
library(bfuncs)
```

```{r}
data("mtcars")
```

Helpful websites: 

[Aggregation with dplyr](https://www.r-bloggers.com/aggregation-with-dplyr-summarise-and-summarise_each/)

[SO summarise with multiple return values](https://stackoverflow.com/questions/22240816/dplyr-summarise-with-multiple-return-values-from-a-single-function)

[R for data science](http://r4ds.had.co.nz/many-models.html)

I originally created the tabstat function awhile back. I didn't create it to work with dplyr pipes and it never worked with grouped variables. Need to address both of those issues below.

I can calculate the mean, and many other statistics for continuous variables using dplyr::summarise. However, dplyr::summarise can only be used to return single number statistics. Because 95% CI's do not fit this requirement, I have to build my own function.

Acutally, maybe I should just use the do function and make a note in r notes. still need to find a function that returns 95% CI's to use with "do". 

# Example 1: Mean no grouping

```{r}
mtcars %>% 
  summarise(mean_mpg = mean(mpg))
```

# Example 2: Mean with grouping

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(mean_mpg = mean(mpg))
```

# Example 3: Tabstat (multiple return values) no grouping

```{r}
mtcars %>% 
  do(tabstat(.$mpg, stats = c("n", "ci")))
```

# Example 4: Tabstat (multiple return values) with grouping

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  do(tabstat(.$mpg, stats = c("n", "ci")))
```

So, using "do" isn't necessarily the most elegant solution ever, but it seems to work just fine.


```{r echo=FALSE}
sessionInfo()
```
